---
title: "final_191892"
author: "Mariia Shekhovtsova"
date: "2024-01-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Assignment task:

Analyzing number of schools
a. Find information about schools for a city you will choose (it can be any city
from anywhere).
i. Lat-long information
ii. Any other information about schools would be appreciated e.g.
number of teachers, students
b. Create a leaflet map and show schools as a point.
c. Have a pop-up menu per each point. Each pop-up should show
i. Name of the school
ii. Name of the city where the school is located.


```{r libraries, include=FALSE}
#install.packages("ggplot2")
#install.packages("ggmap")
#install.packages("maps")
#install.packages("mapdata")
#install.packages("leaflet")
#install.packages("sf")
#install.packages("RColorBrewer")

library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(leaflet)
library(sf)
library(RColorBrewer)
library(mapboxapi)
library(mapdeck)
```

##  49 schools in Calgary, Canada.
For 49 schools of Calgary, Canada, I am going to make a leaflet map and show schools as a point with a pop-up menu, 
where city, school name, level and grade would appear.

Next, I'm going to add isochrone to see from which places it is possible to reach a school by car in 3 minutes.
Such information is important for making good choice and for time planing.

## Creating school data set, Passing mapbox token for isochrone
Data set contains 
1. name, 
2. longitude, 
3. latitude 
and also special parameters: 
4.Level (High school, Middle and Elementary)
5.Max_Grade 

```{r ex1, echo=TRUE, fig.keep='all'}
# School names
school_names <- c(
  "Crescent Heights High School", "Jack James High School", "Lord Beaverbrook High School",
  "Maple Ridge School", "Robert Warren School", "Woodlands School", "Nickle School",
  "McKenzie Lake School", "Altadore School", "Central Memorial High School",
  "Glenbrook School", "Riverbend School", "Sherwood School", "Western Canada High School",
  "Earl Grey School", "Colonel Walker School", "Abbeydale School", "Chris Akkerman School",
  "Douglas Harkness School", "Forest Lawn High School", "Sir Wilfrid Laurier School",
  "Taradale School", "West Dover School", "Radisson Park School", "Lester B. Pearson High School",
  "AADAC", "Alex Munro School", "Beddington Heights School", "Buchanan School",
  "Catherine Nichols Gunn School", "Coventry Hills School", "Alice Jamieson Girls' Academy",
  "Christine Meikle School", "James Fowler High School", "John G. Diefenbaker High School",
  "Senator Patrick Burns School", "William Aberhart High School", "Arbour Lake School",
  "Belvedere Parkway School", "Citadel Park School", "Hawkwood School", "Royal Oak School",
  "Dalhousie School", "Simon Fraser School", "Thomas B. Riley School", "Varsity Acres School",
  "Silver Springs School", "Scenic Acres School", "Dr. Gordon Townsend"
)

# Latitude
latitude <- c(
  51.06171952, 51.05200665, 50.98073203, 50.98521341, 50.9426456, 50.93867223, 50.9411272,
  50.90674808, 51.0140955, 51.00847758, 51.023858, 50.97267535, 50.99405213, 51.03742272,
  51.03130068, 51.03465332, 51.05683801, 51.06415153, 51.07395873, 51.04293213, 51.04599811,
  51.12124767, 51.02633759, 51.04630676, 51.07704003, 51.04366864, 51.12082529, 51.13185073,
  51.08574386, 51.11191215, 51.16553899, 51.06001766, 51.05814405, 51.08685408, 51.11291947,
  51.08857701, 51.07893976, 51.13457405, 51.09313301, 51.14588387, 51.13050944, 51.14561526,
  51.110415, 51.11584488, 51.08790163, 51.0912053, 51.10599196, 51.12033693, 51.07476395
)

# Longitude
longitude <- c(
  -114.0654495, -113.9626091, -114.0621727, -114.0499064, -114.0852237, -114.1077005, -114.0615664,
  -113.9847679, -114.0992716, -114.1138061, -114.1549649, -114.0185411, -114.0117466, -114.0772103,
  -114.0822129, -114.0191304, -113.929612, -113.9635133, -113.9460222, -113.9695208, -113.988828,
  -113.9379524, -113.9906314, -113.9923809, -113.9569991, -114.0770105, -114.0612257, -114.0793795,
  -114.0627841, -114.0579054, -114.0514432, -114.0477123, -114.038277, -114.069975, -114.0702163,
  -114.1153224, -114.1150039, -114.2093393, -114.2112563, -114.1733239, -114.1788549, -114.2224024,
  -114.1496443, -114.1416607, -114.1887857, -114.1473831, -114.1910001, -114.2155934, -114.1482353
)

# School Level
level <- c(
  "High School", "High School", "High School", "Elementary", "Middle", "Elementary", "Middle",
  "Elementary", "Elementary", "High School", "Elementary", "Elementary", "Middle", "High School",
  "Elementary", "Elementary", "Elementary", "Elementary", "Middle", "High School", "Middle",
  "Elementary", "Elementary", "Elementary", "High School", "Special", "Elementary", "Elementary",
  "Elementary", "Elementary", "Elementary", "Elementary", "Middle", "Special", "High School",
  "High School", "Middle", "High School", "Middle", "Middle", "Elementary", "Elementary",
  "Elementary", "Middle", "Middle", "Elementary", "Elementary", "Elementary", "Special"
)

# Grade
max_grade <- c(
  12,12,12,4,9,6,9,4,6,12,6,6,9,12,6,6,6,4,6,12,9,6,6,4,12,12,
  6,6,6,6,4,9,15,12,12,9,12,9,6,4,6,6,4,9,9,6,6,4,15)

# data frame 
schools <- data.frame(
  school = school_names,
  latitude = latitude,
  longitude = longitude,
  level = level,
  max_grade = max_grade
)

calgary_coordinates <- c(51.05011, -114.08529)

# Passing mapbox token
Sys.setenv(MAPBOX_API_KEY = "pk.eyJ1IjoibWFyaWEyMDI0IiwiYSI6ImNscXdsdGZ0dTA1aHUybG1pMnF5cnRwczcifQ.Q_g6RwM_KZr10fbD3uIBVw")
```

## Creating a map

As I want to see a point and an isochrone on one map, I am creating a variable base_map to store data from leaflet.


```{r}
# Create a base map
base_map <- leaflet(data = schools) %>%
  setView(lng = calgary_coordinates[2], lat = calgary_coordinates[1], zoom = 10) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addMarkers(
    lat = ~latitude,
    lng = ~longitude,
    popup = paste0(
      "<b>City: </b>", 'Calgary', "<br>",
      "<b>Name: </b>", schools$school, "<br>",
      "<b>Level: </b>", schools$level, "<br>",
      "<b>Grade: </b>", schools$max_grade, "<br>"
    )
  )
```

## Add isochrone polygons using mapdeck
Here we need a loop, where:

First step is to map the name, lon and lat of each school, because in this case mapdeck doesn't know the location of each school clearly.

Second step is to calculate isochrone for the school

And final one is to add isochrone polygons to the map
```{r}
# A loop through schools to add isochrone polygons
for (i in seq_len(nrow(schools))) {
  school_name <- schools$school[i]
  lon <- schools$longitude[i]
  lat <- schools$latitude[i]
  
  # calculating isochrone for the current school
  isochrone <- mb_isochrone(paste(lon, lat, sep = ","), profile = "driving", time = 3)
  
  # adding isochrone polygons to the map
  base_map <- base_map %>%
    addPolygons(
      data = isochrone,
      fillColor = "violet",
      fillOpacity = 0.25,
      layerId = paste("isochrone_", school_name),
      group = "Isochrones"
    )
}
```
## Option to turn on and off isochrone map and showing the final map

To be able to focus on points only, I can turn off isochrone layer by adding a control.
To see the map I just call the variable base_map.

```{r}
# layer control
base_map <- base_map %>%
  addLayersControl(
    overlayGroups = c("Isochrones"),
    options = layersControlOptions(collapsed = FALSE)
  )

# final map
base_map
```


